Neural Connectome of the Ctenophore Statocyst

### 2025-05-06 manuscript/Manuscript.qmd
conclusion

>>>>

### Conclusion

How do differences in balancer cilia coordination across anatomical planes shape behavior? In *Mnemiopsis*, the animal swims upright with its mouth facing upward, and performs 90-degree rotational turns when hunting prey or responding to disturbances. Large rotational maneuvers may involve balancer cells in the tentacular plane, while sagittal-plane balancers may contribute to postural stabilization—but these remain hypotheses.

Our findings reveal a circuit composed of ANN neurons and bridge cells that coordinates ciliary beating rhythms in space and time without sensory input, and features a recursive architecture in which ciliary outputs are relayed back into the circuit via bridge cells. Such a configuration suggests that even with a limited number of cells, the system is capable of flexible and robust rhythm control.

As Jagar et al. (2010) once described, this organ evokes the metaphor of a “ciliary brain.” The network we describe here similarly orchestrates output through dynamic, self-regulating coordination. These dynamics, visualized through live imaging of the entire cell ensemble, offer a window into the core functions of a nervous system—not in complexity, but in principle.
